Robust estimation of principal components from depth-based multivariate rank covariance matrix

### 2018-03-20 MS/Sinica/SS-template - Copy.tex
first commit in shubhobm

\documentclass[fleqn,11pt]{article}
\usepackage{mycommands,amssymb,amsmath,amsthm,color,pagesize,outlines,cite,caption}
\usepackage[pdftex]{graphicx}
\usepackage[round]{natbib}
\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{0.9in}
\addtolength{\textheight}{0.9in}
\addtolength{\topmargin}{-.4in}

\usepackage{hyperref} % for linking references 
\usepackage{setspace}
\doublespacing

%\DeclareMathOperator*{\vec}{vec}
\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator*{\Tr}{Tr}

@@

\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Lemma}[Theorem]{Lemma}
\newtheorem{Corollary}[Theorem]{Corollary}
\theoremstyle{definition} \newtheorem{Definition}[Theorem]{Definition}

\title{Robust estimation of principal components from depth-based multivariate rank covariance matrix}
\date{}
\author{Subhabrata Majumdar}
\maketitle

Abstract:
Analyzing principal components for multivariate data from its spatial sign covariance matrix (SCM) has been proposed as a computationally simple and robust alternative to normal PCA, but it suffers from poor efficiency properties and is actually inadmissible with respect to the maximum likelihood estimator. Here we use data depth-based spatial ranks in place of spatial signs to obtain the orthogonally equivariant Depth Covariance Matrix (DCM) and use its eigenvector estimates for PCA. We derive asymptotic properties of the sample DCM and influence functions of its eigenvectors. The shapes of these influence functions indicate robustness of estimated principal components, and good efficiency properties compared to the SCM. Finite sample simulation studies show that principal components of the sample DCM are robust with respect to deviations from normality, as well as are more efficient than the SCM and its affine equivariant version, Tyler's shape matrix. Through two real data examples, we also show the effectiveness of DCM-based PCA in analyzing high-dimensional data and outlier detection, and compare it with other methods of robust PCA.
\vspace{.5cm}

Keywords:
Data depth; Principal components analysis; Robustness; Sign covariance matrix; Multivariate ranking

\newpage
\section{Introduction}
In multivariate analysis, the study of principal components is important since it provides a small number of uncorrelated variables from a potentially larger number of variables, so that these new components explain most of the underlying variability in the original data. In case of multivariate normal distribution, the sample covariance matrix provides the most asymptotically efficient estimates of eigenvectors/ principal components, but it is extremely sensitive to outliers as well as relaxations of the normality assumption. To address this issue, several robust estimators of the population covariance or correlation matrix have been proposed which can be used for Principal Components Analysis (PCA). They can be roughly put into these categories: robust, high breakdown point estimators that are computation-intensive \citep{rousseeuw85, maronna76}; M-estimators that are calculated by simple iterative algorithms but do not necessarily possess high breakdown point \citep{huber77, tyler87}; and symmetrised estomators that are highly efficient and robust to deviations from normality, but sensitive to outliers and computationally demanding \citep{dumbgen98, sirkia07}.

@@

The chapter is arranged in the following fashion. Section \ref{section:sec2} provides preliminary theoretical concepts required for developments in the subsequent sections. Section \ref{section:dcmSection} introduces the Depth Covariance Matrix (DCM) and states some basic results related to this. Section \ref{section:sec4} provides asymptotic results regarding the sample DCM, calculated using data depths with respect to the empirical distribution function, as well as its eigenvectors and eigenvalues. Section \ref{section:simSection} focuses solely on principal component estimation using the sample DCM. We obtain influence functions and asymptotic efficiencies for eigenvectors of the DCM. We also compare their finite sample efficiencies for several multinormal and multivariate $t$-distributions with those of the SCM, Tyler's scatter matrix and its depth-weighted version through a simulation study. Section \ref{section:sec6} presents two applications of the methods we develop on real data. Finally, we wrap up our discussion in Section \ref{section:sec7} by giving a summary of our findings and providing some potential future areas of research. Appendices \ref{section:appA} and \ref{section:appB} contain all technical details and proofs of the results we derive.

\section{Preliminaries}\label{section:sec2}
\subsection{Spatial signs and sign covariance matrix}
Given a vector $\bfx \in \mathbb{R}^p$, its spatial sign is defined as the vector valued function \citep{MottonenOja95}:
%
$$ \bfS(\bfx) = \begin{cases} \bfx\| \bfx \|^{-1} \quad \mbox{if }\bfx \neq \bf0\\
\bf0 \quad \mbox{if }\bfx = \bf0 \end{cases} $$
%
When $\bfx$ is a random vector that follows an elliptic distribution $|\Sigma|^{-1/2} f((\bfx - \bfmu)^T \Sigma^{-1} (\bfx - \bfmu))$, with a mean vector $\bfmu$ and covariance matrix $\Sigma$, the sign vectors $\bfS(\bfx - \bfmu)$ reside on the surface of a $p$-dimensional unit ball centered at $\bfmu$. Denote by $\Sigma_S(\bfX) = E\bfS (\bfX - \bfmu)\bfS (\bfX - \bfmu)^T$ the covariance matrix of spatial signs, or the \textit{Sign Covariance Matrix} (SCM). The transformation $\bfx \mapsto \bfS(\bfx - \bfmu)$ keeps eigenvectors of population covariance matrix unchanged, and eigenvectors of the sample SCM $ \hat \Sigma_S = \sum_{i=1}^n \bfS (\bfx_i - \bfmu)\bfS (\bfx_i - \bfmu)^T/n $ are $\sqrt n$-consistent estimators of their population counterparts \citep{taskinen12}.

@@

\subsection{Data depth and outlyingness}
For any multivariate distribution $F = F_\bfX$ belonging to a set of distributions $\mathcal F$, the depth of a point $\bfx \in \mathbb{R}^p$, say $D(\bfx, F_\bfX)$ is any real-valued function that provides a 'center outward ordering' of $\bfx$ with respect to $F$ \citep{zuo00}. \cite{liu90} outlines the desirable properties of a statistical depth function:

\vspace{1em}
\noindent\textbf{(D1)} \textit{Affine invariance}: $D(A\bfx + \bfb, F_{A\bfX+\bfb}) = D(\bfx, F_\bfX)$;

\noindent\textbf{(D2)} \textit{Maximality at center}: $D(\bftheta, F_\bfX) = \sup_{\bfx\in \mathbb{R}^p} D(\bfx, F_\bfX)$ for $F_\bfX$ having center of symmetry $\bftheta$. This point is called the \textit{deepest point} of the distribution.;

\noindent\textbf{(D3)} \textit{Monotonicity with respect to deepest point}: $D(\bfx; F_\bfX) \leq D(\bftheta + a(\bfx - \bftheta), F_\bfX)$, $\bftheta$ being deepest point of $F_\bfX$.;

\noindent\textbf{(D4)} \textit{Vanishing at infinity}: $D(\bfx; F_\bfX) \rightarrow 0$ as $\|\bfx\| \rightarrow \infty $.
\vspace{1em}

In (D2) the types of symmetry considered can be central symmetry, angular symmetry and halfspace symmetry. Also for multimodal probability distributions, i.e. distributions with multiple local maxima in their probability density functions, properties (D2) and (D3) are actually restrictive towards the formulation of a reasonable depth function that captures the shape of the data cloud. In our derivations that follow, we replace these two by a weaker condition:

\vspace{1em}
\noindent\textbf{(D2*)} \textit{Existence of a maximal point}: The maximum depth over all distributions $F$ and points $\bfx$ is bounded above, i.e. $ \sup_{F_\bfX \in \mathcal F} \sup_{\bfx\in \mathbb{R}^p} D(\bfx, F_\bfX) < \infty $. We denote this point by $M_D(F_\bfX)$.
\vspace{1em}

A real-valued function measuring the outlyingness of a point with respect to the data cloud can be seen as the opposite of what data depth does. Indeed, such functions have been used to define several depth functions, for example simplicial depth, projection depth and $L_p$-depth. Keeping with the spirit of the utility of these functions we name them `htped': literally the reverse of `depth', and give a general definition of such functions as a transformation on any depth function:

@@

\item \textbf{Mahalanobis depth} (MhD) \citep{LiuPareliusSingh99} is based on the Mahalanobis distance of $\bfx$ to $\bfmu$ with respect to $\Sigma$: $d_\Sigma(\bfx, \bfmu) = \sqrt{(\bfx - \bfmu)^T \Sigma^{-1} (\bfx - \bfmu)}$. It is defined as
%
$$ MhD(\bfX, F) = \frac{1}{1 + d^2_\Sigma (\bfx - \bfmu)} $$
%
note here that $d_\Sigma(\bfx,\bfmu)$ can be seen as a valid htped function of $\bfx$ with respect to $F$.

\item \textbf{Projection depth} (PD) \citep{zuo03} is another depth function based on an outlyingness function. Here that function is
%
$$ O(\bfx, F) = \sup_{\| \bfu \| = 1} \frac{| \bfu^T\bfx - m(\bfu^T\bfX)|}{s(\bfu^T\bfX)} $$
%
where $m$ and $s$ are some univariate measures location and scale, respectively. Given this the depth at $\bfx$ is defined as $PD(\bfx, F) = 1/(1+O(\bfx, F))$.
\end{itemize}

@@

\section{Depth-based rank covariance matrix} \label{section:dcmSection}

Consider a vector-valued random variable $\bfX \in \mathbb{R}^p$ . Data depth is as much a property of the random variable as it is of the underlying distribution, so for ease of notation while working with transformed random variables, from now on we shall be using $D_\bfX(\bfx) = D(\bfx, F)$ to denote the depth of a point $\bfx$. Now, given a depth function $D_{\bfX}(\bfx)$ (equivalently, an htped function $\tilde D_\bfX(\bfx) = \tilde D(\bfx, F)$), transform the original random variable as: $\tilde \bfx = \tilde D_\bfX(\bfx) \bfS(\bfx - \bfmu)$, $\bfS(.)$ being the spatial sign functional. The transformed random variable $\tilde \bfX$ can be seen as the multivariate rank corresponding to $\bfX$ (e.g. \cite{serfling2006}). The notion of multivariate ranks goes back to \cite{PuriSenBook}, where they take the vector consisting of marginal univariate ranks as multivariate rank vector. Subsequent definitions of multivariate ranks were proposed by \cite{MottonenOja95,HallinPaindaveine02} and \cite{Chernozhukov14}. Compared to these formulations, our definition of multivariate ranks works for any general depth function, and provides an intuitive extension to any spatial sign-based methodology.

Figure \ref{fig:rankplot} gives an idea of how the multivariate rank vector $\tilde \bfX$ is distributed when $\bfX$ has a bivariate normal distribution. Compared to the spatial sign, which are distributed on the surface of $p$-dimensional unit ball centered at $\bfmu$, these spatial ranks have the same direction as original data and reside \textit{inside} the $p$-dimensional ball around $\bfmu$ that has radius $M_D(F)$ (which, for the case of halfspace depth, equals 0.5).

\begin{figure}[t]
	\captionsetup{singlelinecheck=off}
	\centering
		\includegraphics[height=6cm]{../Codes/ranks}
	\caption{(Left) 1000 points randomly drawn from $\mathcal N_2\left((0,0)^T, \left(\protect\begin{smallmatrix} 5 & -4 \\ -4 & 5 \protect\end{smallmatrix}\right)\right) $ and (Right) their multivariate ranks based on halfspace depth}
	\label{fig:rankplot}
\end{figure}

@@

Unlike sign covariance matrices and symmetrized sign covariance matrices \citep{dumbgen98}, however, attempting to derive an affine equivariant counterpart (as opposed to only orthogonal equivariance) of the DCM through an iterative approach analogous to \cite{tyler87} will not result in anything new. This is because Tyler's scatter matrix $\Sigma_T$ is defined as the implicit solution to the following equation:
\begin{equation} \label{equation:tylerEq}
\Sigma_T = \mathbb E\left[ \frac{(\bfx - \bfmu) (\bfx - \bfmu)^T}{(\bfx - \bfmu)^T \Sigma_T^{-1} (\bfx - \bfmu)} \right]
\end{equation}
and simply replacing $\bfx$ by its multivariate rank counterpart $\tilde\bfx$ will not change the estimate $\Sigma_T$ as $\bfx$ and $\tilde \bfx$ have the same directions. Instead we consider a depth-weighted version of Tyler's scatter matrix (i.e. weights $(\tilde D_\bfX(\bfx))^2$ in right side of (\ref{equation:tylerEq})) in the simulations in Section \ref{section:simSection}. The simulations show that it has slightly better finite-sample efficiency than $\Sigma_T$ but has same asymptotic performance. We conjecture that its concentration properties can be obtained by taking an approach similar to \cite{soloveychik14}.

\section{Asymptotic results}\label{section:sec4}
\subsection{The sample DCM}
Let us now consider $n$ iid random draws from our elliptic distribution $F$, say $\bfX_1,...,\bfX_n$. For ease of notation, denote $SS(\bfx; \bfmu) = \bfS(\bfx - \bfmu) \bfS(\bfx - \bfmu)^T$. Then, given the depth function and known location center $\bfmu$, one can show that the vectorized form of $\sqrt n$-times the sample DCM: $\sum_{i=1}^n (\tilde D_\bfX(\bfx_i))^2SS(\bfx_i; \bfmu) /\sqrt{n}$  has an asymptotic multivariate normal distribution with mean $\sqrt n.vec(E[( (\tilde D_\bfX (\bfX))^2 SS(\bfx; \bfmu)])$ and a certain covariance matrix by straightforward application of the central limit theorem (CLT). But in practice the population depth function $D_\bfX(\bfx) = D(\bfx, F)$ is estimated by the depth function based on the empirical distribution function, $F_n$. Denote this sample depth by $D^n_\bfX (\bfx) = D(\bfx, F_n)$. Here we make the following assumption regarding its relation to $D_\bfX(\bfx)$:

\vspace{1em}
\noindent\textbf{(D5)} \textit{Uniform convergence}: $\sup_{\bfx \in \mathbb R^p} | D^n_\bfX (\bfx) - D_\bfX (\bfx) | \rightarrow 0$ as $n \rightarrow \infty $.
\vspace{1em}

The assumption that empirical depths converge uniformly at all points $\bfx$ to their population versions holds under very mild conditions for several well known depth functions: for example projection depth \citep{zuo03} and simplicial depth \citep{Dumbgen92}. One also needs to replace the known location parameter $\bfmu$ by some estimator $\hat\bfmu_n$. Examples of robust estimators of location that are relevant here include the spatial median \citep{haldane48,brown83}, Oja median \citep{oja83}, projection median \citep{zuo03} etc. Now, given $D^n_\bfX(.)$ and $\hat \bfmu_n$, to plug them into the sample DCM and still go through with the CLT we need the following result:

\begin{Lemma} \label{Lemma:lemma1}
Consider a random variable $\bfX \in \mathbb{R}^p$ having a continuous and symmetric distribution with location center $\bfmu$ such that $E\|\bfx - \bfmu \|^{-3/2} < \infty$. Given $n$ random samples from this distribution, suppose $\hat\bfmu_n$ is an estimator of $\bfmu$ so that $\sqrt n (\hat\bfmu_n - \bfmu) = O_P(1) $. Then with the above notations, and given the assumption (D5) we have
%
$$ \sqrt n \left[
\frac{1}{n} \sum_{i=1}^n (\tilde D^n_\bfX (\bfx_i))^2 SS(\bfx_i; \hat\bfmu_n) -
\frac{1}{n} \sum_{i=1}^n (\tilde D_\bfX (\bfx_i))^2 SS(\bfx_i; \bfmu) \right]
\stackrel{P}{\rightarrow} 0 $$
\end{Lemma}

We are now in a position to state the result for consistency of the sample DCM:

>>>>

\documentclass[10pt]{book}
\usepackage[sectionbib]{natbib}
\usepackage{array,epsfig,fancyheadings,rotating,caption}
\usepackage{hyperref}
\usepackage{mycommands}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textwidth=31.9pc
\textheight=46.5pc
\oddsidemargin=1pc
\evensidemargin=1pc
\headsep=15pt
%\headheight=.2cm
\topmargin=.6cm
\parindent=1.7pc
\parskip=0pt

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage{amsthm}

\setcounter{page}{1}
\pagestyle{fancy}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{fancy}
\def\n{\noindent}
\lhead[\fancyplain{} \leftmark]{}
\chead[]{}
\rhead[]{\fancyplain{}\rightmark}
\cfoot{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Lemma}[Theorem]{Lemma}
\newtheorem{Corollary}[Theorem]{Corollary}
\theoremstyle{definition} \newtheorem{Definition}[Theorem]{Definition}
%\newtheorem{proof}{Proof}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\renewcommand{\baselinestretch}{1.2}

\markright{ \hbox{\footnotesize\rm Statistica Sinica
%{\footnotesize\bf 24} (201?), 000-000
}\hfill\\[-13pt]
\hbox{\footnotesize\rm
%\href{http://dx.doi.org/10.5705/ss.20??.???}{doi:http://dx.doi.org/10.5705/ss.20??.???}
}\hfill }

\markboth{\hfill{\footnotesize\rm FIRSTNAME1 LASTNAME1 AND FIRSTNAME2 LASTNAME2} \hfill}
{\hfill {\footnotesize\rm FILL IN A SHORT RUNNING TITLE} \hfill}

\renewcommand{\thefootnote}{}
$\ $\par

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\fontsize{10.95}{14pt plus.8pt minus .6pt}\selectfont
\vspace{0.8pc}
\centerline{\large\bf Robust estimation of principal components}
\vspace{2pt}
\centerline{\large\bf from depth-based multivariate rank covariance matrix}
\vspace{.4cm}
\centerline{Author(s)}
\vspace{.4cm}
\centerline{\it Affiliation(s)}
\vspace{.55cm}
\fontsize{9}{11.5pt plus.8pt minus .6pt}\selectfont

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{quotation}
\noindent {\it Abstract:}
Analyzing principal components for multivariate data from its spatial sign covariance matrix (SCM) has been proposed as a computationally simple and robust alternative to normal PCA, but it suffers from poor efficiency properties and is actually inadmissible with respect to the maximum likelihood estimator. Here we use data depth-based spatial ranks in place of spatial signs to obtain the orthogonally equivariant Depth Covariance Matrix (DCM) and use its eigenvector estimates for PCA. We derive asymptotic properties of the sample DCM and influence functions of its eigenvectors. The shapes of these influence functions indicate robustness of estimated principal components, and good efficiency properties compared to the SCM. Finite sample simulation studies show that principal components of the sample DCM are robust with respect to deviations from normality, as well as are more efficient than the SCM and its affine equivariant version, Tyler's shape matrix. Through two real data examples, we also show the effectiveness of DCM-based PCA in analyzing high-dimensional data and outlier detection, and compare it with other methods of robust PCA.
\par

\vspace{9pt}
\noindent {\it Key words and phrases:}
Data depth; Principal components analysis; Robustness; Sign covariance matrix; Multivariate ranking
\par
\end{quotation}\par

\def\thefigure{\arabic{figure}}
\def\thetable{\arabic{table}}

\fontsize{10.95}{14pt plus.8pt minus .6pt}\selectfont

\setcounter{chapter}{1}
\setcounter{equation}{0} %-1
\noindent {\bf 1. Introduction}

In multivariate analysis, the study of principal components is important since it provides a small number of uncorrelated variables from a potentially larger number of variables, so that these new components explain most of the underlying variability in the original data. In case of multivariate normal distribution, the sample covariance matrix provides the most asymptotically efficient estimates of eigenvectors/ principal components, but it is extremely sensitive to outliers as well as relaxations of the normality assumption. To address this issue, several robust estimators of the population covariance or correlation matrix have been proposed which can be used for Principal Components Analysis (PCA). They can be roughly put into these categories: robust, high breakdown point estimators that are computation-intensive \citep{rousseeuw85, maronna76}; M-estimators that are calculated by simple iterative algorithms but do not necessarily possess high breakdown point \citep{huber77, tyler87}; and symmetrised estomators that are highly efficient and robust to deviations from normality, but sensitive to outliers and computationally demanding \citep{dumbgen98, sirkia07}.

@@

The paper is arranged in the following fashion. Section 2 provides preliminary theoretical concepts required for developments in the subsequent sections. Section 3 introduces the Depth Covariance Matrix (DCM) and states some basic results related to this. Section 4 provides asymptotic results regarding the sample DCM, calculated using data depths with respect to the empirical distribution function, as well as its eigenvectors and eigenvalues. Section 5 focuses solely on principal component estimation using the sample DCM. We obtain influence functions and asymptotic efficiencies for eigenvectors of the DCM. We also compare their finite sample efficiencies for several multinormal and multivariate $t$-distributions with those of the SCM, Tyler's scatter matrix and its depth-weighted version through a simulation study. Section 6 presents two applications of the methods we develop on real data. Finally, we wrap up our discussion in Section 7 by giving a summary of our findings and providing some potential future areas of research. Proofs of all the results we derive are given in the appendix.
\par

\vspace{1em}
\setcounter{chapter}{2}
\setcounter{equation}{0} %-1
\noindent {\bf 2. Preliminaries}

Given a vector $\bfx \in \mathbb{R}^p$, its spatial sign is defined as the vector valued function \citep{locantore99}:
%
$$ \bfS(\bfx) = \begin{cases} \bfx\| \bfx \|^{-1} \quad \mbox{if }\bfx \neq \bf0\\
\bf0 \quad \mbox{if }\bfx = \bf0 \end{cases} $$
%
When $\bfx$ is a random vector that follows an elliptic distribution $|\Sigma|^{-1/2} f((\bfx - \bfmu)^T \Sigma^{-1} (\bfx - \bfmu))$, with a mean vector $\bfmu$ and covariance matrix $\Sigma$, the sign vectors $\bfS(\bfx - \bfmu)$ reside on the surface of a $p$-dimensional unit ball centered at $\bfmu$. Denote by $\Sigma_S(\bfX) = E\bfS (\bfX - \bfmu)\bfS (\bfX - \bfmu)^T$ the covariance matrix of spatial signs, or the \textit{Sign Covariance Matrix} (SCM). The transformation $\bfx \mapsto \bfS(\bfx - \bfmu)$ keeps eigenvectors of population covariance matrix unchanged, and eigenvectors of the sample SCM $ \hat \Sigma_S = \sum_{i=1}^n \bfS (\bfx_i - \bfmu)\bfS (\bfx_i - \bfmu)^T/n $ are $\sqrt n$-consistent estimators of their population counterparts \citep{taskinen12}.

@@

\vspace{.5em}
\noindent {\bf 2.1 Data depth and outlyingness}

For any multivariate distribution $F = F_\bfX$, the depth of a point $\bfx \in \mathbb{R}^p$, say $D(\bfx, F_\bfX)$ is any real-valued function that provides a 'center outward ordering' of $\bfx$ with respect to $F$ \citep{zuo00}. \cite{liu90} outlines the desirable properties of a statistical depth function:

\noindent
(P1) Affine invariance: $D(A\bfx + \bfb, F_{A\bfX+\bfb}) = D(\bfx, F_\bfX)$\\
(P2) Maximality at center: $D(\bftheta, F_\bfX) = \sup_{\bfx\in \mathbb{R}^p} D(\bfx, F_\bfX)$ for $F_\bfX$ having center of symmetry $\bftheta$. This point is called the \textit{deepest point} of the distribution.\\
(P3) Monotonicity with respect to deepest point: $D(\bfx; F_\bfX) \leq D(\bftheta + a(\bfx - \bftheta), F_\bfX)$, $\bftheta$ being deepest point of $F_\bfX$.\\
(P4) Vanishing at infinity: $D(\bfx; F_\bfX) \rightarrow \bf0$ as $\|\bfx\| \rightarrow \infty $.

In (P2) the types of symmetry considered can be central symmetry, angular symmetry and halfspace symmetry. Also for multimodal probability distributions, i.e. distributions with multiple local maxima in their probability density functions, properties (P2) and (P3) are actually restrictive towards the formulation of a reasonable depth function that captures the shape of the data cloud. In our derivations that follow, we replace these two by a slightly weaker condition: (P2*) The \textit{existence} of a maximal point, i.e. $ \sup_{\bfx\in \mathbb{R}^p} D(\bfx, F_\bfX) < \infty $. We denote this point by $M_D(F)$.

A real-valued function measuring the outlyingness of a point with respect to the data cloud can be seen as the opposite of what data depth does. Indeed, such functions have been used to define several depth functions, for example simplicial volume depth, projection depth and $L_p$-depth. Here we give a general definition of such functions as a transformation on any depth function:

@@

\item \textbf{Mahalanobis depth} (MhD) \citep{LiuPareliusSingh99} is based on the Mahalanobis distance of $\bfx$ to $\bfmu$ with respect to $\Sigma$: $d_\Sigma(\bfx, \bfmu) = \sqrt{(\bfx - \bfmu)^T \Sigma^{-1} (\bfx - \bfmu)}$. It is defined as

$$ MhD(\bfX, F) = \frac{1}{1 + d^2_\Sigma (\bfx - \bfmu)} $$

note here that $d_\Sigma(\bfx,\bfmu)$ can be seen as a valid htped function of $\bfx$ with respect to $F$.

\item \textbf{Projection depth} (PD) \citep{zuo03} is another depth function based on an outlyingness function. Here that function is

$$ O(\bfx, F) = \sup_{\| \bfu \| = 1} \frac{| \bfu^T\bfx - m(\bfu^T\bfX)|}{s(\bfu^T\bfX)} $$

where $m$ and $s$ are some univariate measures location and scale, respectively. Given this the depth at $\bfx$ is defined as $PD(\bfx, F) = 1/(1+O(\bfx, F))$.
\end{itemize}

@@

\vspace{1em}
\setcounter{chapter}{3}
\setcounter{equation}{0} %-1
\noindent {\bf 3. Depth-based rank covariance matrix}\vspace{1em}

Consider a vector-valued random variable $\bfX \in \mathbb{R}^p$. Data depth is as much a property of the random variable as it is of the underlying distribution, so for ease of notation while working with transformed random variables, from now on we shall be using $D_\bfX(\bfx) = D(\bfx, F)$ to denote the depth of a point $\bfx$. Now, given a depth function $D_{\bfX}(\bfx)$ (equivalently, an htped function $\tilde D_\bfX(\bfx) = \tilde D(\bfx, F)$), transform the original random variable as: $\tilde \bfx = \tilde D_\bfX(\bfx) \bfS(\bfx - \bfmu)$, $\bfS(.)$ being the spatial sign functional. The transformed random variable $\tilde \bfX$ can be seen as the multivariate rank corresponding to $\bfX$ (e.g. \cite{serfling2006}). Figure \ref{fig:rankplot} gives an idea of how $\tilde \bfX$ is distributed when $\bfX$ has a bivariate normal distribution. Compared to the spatial sign, which are distributed on the surface of $p$-dimensional unit ball centered at $\bfmu$, these spatial ranks have the same direction as original data and reside \textit{inside} the $p$-dimensional ball around $\bfmu$ that has radius $M_D(F)$ (which, for the case of halfspace depth, equals 0.5).

\begin{figure}[t]
	\captionsetup{singlelinecheck=off}
	\centering
		\includegraphics[height=6cm]{ranks}
	\caption{(Left) 1000 points randomly drawn from $\mathcal N_2\left((0,0)^T, \left(\protect\begin{smallmatrix} 5 & -4 \\ -4 & 5 \protect\end{smallmatrix}\right)\right) $ and (Right) their multivariate ranks based on halfspace depth}
	\label{fig:rankplot}
\end{figure}

@@

Unlike sign covariance matrices and symmetrized sign covariance matrices \citep{dumbgen98}, however, attempting to derive an affine equivariant counterpart (as opposed to only orthogonal equivariance) of the DCM through an iterative approach analogous to \cite{tyler87} will not result in anything new. This is because Tyler's scatter matrix $\Sigma_T$ is defined as the implicit solution to the following equation:
%
\begin{equation} \label{equation:tylerEq}
\Sigma_T = \mathbb E\left[ \frac{(\bfx - \bfmu) (\bfx - \bfmu)^T}{(\bfx - \bfmu)^T \Sigma_T^{-1} (\bfx - \bfmu)} \right]
\end{equation}
%
and simply replacing $\bfx$ by its multivariate rank counterpart $\tilde\bfx$ will not change the estimate $\Sigma_T$ as $\bfx$ and $\tilde \bfx$ have the same directions. Instead we consider a depth-weighted version of Tyler's scatter matrix (i.e. weights $(\tilde D_\bfX(\bfx))^2$ in right side of (\ref{equation:tylerEq})) in the simulations in Section \ref{section:simSection}. The simulations show that it has slightly better finite-sample efficiency than $\Sigma_T$ but has same asymptotic performance. We conjecture that its concentration properties can be obtained by taking an approach similar to \cite{soloveychik14}.
%\begin{figure}[h!]
%\includegraphics [angle=-90, scale=0.45]{d:/sinicas/simuh3.eps}\par
%\centerline{\epsfig{file=d:/sinicas/simuh3.eps,angle=-90,width=4.5in}}\par
%\centerline{\epsfig{file=d:/sinicas/simuh4.eps,width=4.5in}}\par

%\caption{Caption of the figure.}
%\end{figure}

\begin{equation}\label{2.1}
\mbox {The 1st display equation of the second section.}
\end{equation}
\begin{equation}\label{2.2}
\mbox {The 2nd display equation of the second section.}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\lhead[\footnotesize\thepage\fancyplain{}\leftmark]{}\rhead[]{\fancyplain{}\rightmark\footnotesize\thepage}%Put this line in Page 2

\setcounter{chapter}{3}
\setcounter{equation}{0} %-1
\noindent {\bf 3. The Third Section}

Contents of the third section.

Contents of the third section.

Contents of the third section.

Contents of the third section.

Contents of the third section.
\begin{equation}\label{3.1}
\mbox {The 1st display equation of the third section.}
\end{equation}
\begin{equation}\label{3.2}
\mbox {The 2nd display equation of the third section.}
\end{equation}
\newpage
\begin{table}
\caption{Caption of the table.}
\label{tab:simulation}\par
\vskip .2cm
\centerline{\tabcolsep=5truept\begin{tabular}{|rcrrrrrrrrr|} \hline
\multicolumn{2}{|c}{Bandwidth}
         & \multicolumn{3}{c}{$h=3$}
         & \multicolumn{3}{c}{$h=4$}
         & \multicolumn{3}{c|}{$h=5$} \\ \hline
         &&&&&&&&&&\\[-11pt]
\multicolumn{2}{|c}{Estimates}
          & \multicolumn{1}{c}{$\hat N_H^{LC}$} & \multicolumn{1}{c}{$\hat N_{\bar H}^{LC}$} & \multicolumn{1}{c}{$\hat N_{\bar H}^{LL}$}
          & \multicolumn{1}{c}{$\hat N_H^{LC}$} & \multicolumn{1}{c}{$\hat N_{\bar H}^{LC}$} & \multicolumn{1}{c}{$\hat N_{\bar H}^{LL}$}
          & \multicolumn{1}{c}{$\hat N_H^{LC}$} & \multicolumn{1}{c}{$\hat N_{\bar H}^{LC}$} & \multicolumn{1}{c|}{$\hat N_{\bar H}^{LL}$} \\[3pt] \hline
$beta(10,10)$ & BIAS  &-22.5&-14.8&14.0&-13.3&-6.9&12.5&-8.2&-4.7&11.5 \\
$\bar{p}=0.500$ & S.E. &13.8&14.9&12.1&12.6&14.6&11.7&15.1&19.3&15.5 \\
$cv=0.218$ & RMSE &26.4&21.0&18.5&18.3&16.2&17.1&17.2&19.9&19.3
\\ \hline
$beta(5,5)$ & BIAS &-32.2&-19.9&5.4&-21.9&-11.8&4.9&-15.6&-8.4&5.1 \\
$\bar{p}=0.500$ & S.E. &15.8&16.2&12.3&14.3&19.8&12.9&16.4&19.8&14.6 \\
$cv=0.302$ & RMSE &35.9&25.7&13.4&26.1&15.9&11.9&22.6&21.5&15.4
\\ \hline
$beta(4,8)$ & BIAS &-53.7&-29.1&-10.8&-42.0&-19.3&-8.4&-34.4&-13.9&-7.6 \\
$\bar{p}=0.333$ & S.E. &20.8&19.4&18.1&18.5&18.8&15.6&19.2&20.6&16.4 \\
$cv=0.392$ & RMSE &57.6&34.9&21.1&45.9&26.9&17.8&39.4&24.9&18.1
\\ \hline
$beta(3,5)$ & BIAS &-57.4&-32.4&-15.5&-45.6&-22.8&-13.1&-37.4&-17.1&-11.4 \\
$\bar{p}=0.375$ & S.E. &21.2&19.9&17.1&18.9&18.8&15.0&19.6&21.6&15.6 \\
$cv=0.430$ & RMSE &61.2&38.1&23.1&49.4&29.5&19.9&42.3&27.5&19.3
\\ \hline
\end{tabular}}
\end{table}
